---
title: "Playing Style"
author: "Braden Churcher"
date: "27 June 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load("~/Documents/UNI/Winter Research Project/Data/atp_serves.RData")

library(ggplot2)
library(dplyr)

library(purrr)


load("atp_serves.RData")


g <- function(coef, t) coef %*% t^(0:3)
g(unlist(atp_serves[1,c("raw.x0.1","raw.x1.1","raw.x2.1","raw.x3.1")]),atp_serves$start.1[1])

fedvmurray <- atp_serves[atp_serves$matchid=="AO2016F_Djokovic_Murray",]

table(atp_serves$server)
table(atp_serves$receiver)

```

## R Markdown
*Brain storm of serve styles*
List of Serve Characteristics for feature selection:
1st vs 2nd Serve
Points in game? if reciever is on 40 or server is on 40. Does this differ? are there risk takers?
Winning serves
Receivers Hand (R/L)
Riskiness and 
Types and key points:
Kick Serve:
  - Height of bounce (Z at )
      - Where was the Ball received?
  - Height over the net
  - Location ball is struck (is this different? To get top spin is it in a different spot)
  - Second serve characteristics
  - Speed
  - Spin? 
    - squared coefficients c2 --generate points function
    - How do we calculate this?
  - Location in Service Box

Down the tee (Flat vs Spin/Kick)
  - Ball location in service box, depth looks important
  - Speed
  - Location ball is hit. For right handers in the ad court this could be back past the centre line on the      deuce side
Outwide (Flat vs Spin/Kick)

Variables needed for Clustering:
Player - Server
X, Z, Y From start to finish.
Coefficients 
Speed
Receiver
Volicity 
Spin
Serve
Serve classification


** Random Plots ***  
## Roger Federer Receiving ##
```{r}

fedr <- atp_serves %>% filter(receiver == "FEDERER")
ggplot(fedr, aes(x=center.x, y=center.y, color=factor(serve_classification))) + geom_point() + facet_wrap(~side)

```
## Serve location by player ##
```{r}
ggplot(atp_serves_sub, aes(x=start.y, y=start.z, color=factor(server))) + geom_point()
```

## Most Aces ##

```{r}
table(atp_serves$server, atp_serves$serve_classification== "0")
```
## Raonic Aces ##
``` {r}
Rao<- atp_serves %>% filter(server == "RAONIC")
ggplot(Rao, aes(x=start.y, y=start.z, color=factor(serve_classification == "0"))) + geom_point()
ggplot(Rao, aes(x=start.x, y=start.z, color=factor(serve_classification == "0"))) + geom_point() 
ggplot(Rao, aes(x=center.x, y=center.y, color=factor(serve_classification))) + geom_point() 
```

``` {r}
# Net Clearance Height by 1st and second serve #
#work out how to solve for x in arc one at x=0 for all serves#
ggplot(atp_serves, aes(x=center.x, y=net.clearance, color=factor(serve.x))) + geom_point() 
ggplot(atp_serves, aes(x=center.y, y=net.clearance, color=factor(serve.x))) + geom_point() 
```
Using the distance at which the ball bounced in the x direction, you can see the difference in serves. The first serve being clearly longer and slightly lower over the net. 
Using the width at which the ball bounced in the court you can see the how second serves are generally much safer. The height over the net though may not be exactly easy to determine if it is different, there does seem to be less serves at the lower end of the scale and more variation in the second serve. 
Interestingly you can see how the net slouches in the the middle of the court. 

```{r}
# Different points in games #
ggplot(atp_serves, aes(x=center.x, y=center.y, color=factor(serve.x))) + geom_point() + facet_wrap(~receiver_score)

```


```{r}
# Serve classification by first and second serve #
table(atp_serves$serve_classification, atp_serves$serve.x)
```

Note that there are 37 double faults in the data, 6 aces on the second serve and when filtering by first serve for clustering, 357 serves will be lost. 

```{r}
## Roger Federer Receiving ##
fedr <- atp_serves %>% filter(receiver == "FEDERER")
ggplot(fedr, aes(x=center.x, y=center.y, color=factor(serve_classification))) + geom_point() + facet_wrap(~side)

## Most Aces ##
table(atp_serves$server, atp_serves$serve_classification== "0")

## Raonic Aces ##
Rao<- atp_serves %>% filter(server == "RAONIC")
ggplot(Rao, aes(x=start.y, y=start.z, color=factor(serve_classification == "0"))) + geom_point()
```


You can also embed plots, for example:

```{r pressure, echo=FALSE}
### Mitchell code for Court Position of Ball ###
## Legacy generate points function ##
traj_coords <- function(start, dir, flip, c0, c1, c2, c3, tm=0) { 
  tm <- tm + start 
  if (dir != "z") { 
    c0 <- flip * c0 
    c1 <- flip * c1 
    c2 <- flip * c2 
    c3 <- flip * c3 
  } 
  pos <- c0 + c1*tm + c2*tm^2 + c3*tm^3 
  return(pos) 
} 
 
GeneratePoints <- function(data, ..., arc1=10, arc3=0, plot=FALSE){ 
  require(tidyr) 
  require(dplyr) 
  require(purrr) 
   
  #arc2*(start.3 - (start.1+duration.arc1)) 
  data <- data %>% mutate(duration = arc1*duration.arc1 + arc3*duration.arc3, flip=sign(raw.x0.1)) 
   
  dots <- c("serveid", sapply(substitute(list(...))[-1], deparse)) 
  extravars <- match(dots[dots%in%colnames(data)], colnames(data)) 
   
  out <- data.frame() 
   
  if(arc1!=0){ 
    arc1p <- data %>% select(extravars, start.1, duration.arc1, flip, raw.x0.1:raw.z3.1) %>% 
      gather(coef, value, -(1:length(extravars)), -start.1, -duration.arc1, -flip) %>% 
      separate(coef, c("junk1", "coef", "junk2"), sep="\\.") %>% 
      select(-junk1, -junk2) %>% 
      mutate(dir=substr(coef, 1, 1), coef=substr(coef, 2, 2)) %>% 
      spread(coef, value) %>% 
      rename("start" = `start.1`, "duration"=`duration.arc1`, "c0"=`0`, "c1"=`1`, "c2"=`2`, "c3"=`3`) 
      arc1points <- seq(0,1,length.out=arc1) %>% 
        map_df(function(x) mutate(rowwise(arc1p), time=start+duration*x, pos=traj_coords(start, dir, flip, c0, c1, c2, c3, tm=duration*x))) 
      out <- rbind(out, cbind(arc=1, arc1points)) 
  } 
   
  if(arc3!=0){ 
    arc3p <- data %>% select(extravars, start.3, duration.arc3, flip, raw.x0.3:raw.z3.3) %>% 
      gather(coef, value, -(1:length(extravars)), -start.3, -duration.arc3, -flip) %>% 
      separate(coef, c("junk1", "coef", "junk2"), sep="\\.") %>% 
      select(-junk1, -junk2) %>% 
      mutate(dir=substr(coef, 1, 1), coef=substr(coef, 2, 2)) %>% 
      spread(coef, value) %>% 
      rename("start" = `start.3`, "duration"=`duration.arc3`, "c0"=`0`, "c1"=`1`, "c2"=`2`, "c3"=`3`) 
    arc3points <- seq(0,1, length.out=arc3) %>% 
      map_df(function(x) mutate(rowwise(arc3p), time=start+duration*x, pos=traj_coords(start, dir, flip, c0, c1, c2, c3, tm=duration*x))) 
    out <- rbind(out, cbind(arc=3, arc3points)) 
  } 
   
  if(plot){ 
    extravars <- match(dots[dots%in%colnames(out)], colnames(out)) 
    out <- out %>% select(arc, extravars, flip, dir, time, pos) %>% 
      spread(dir, pos) 
  } 
   
  return(out) 
} 
 
#DEMO 
 
sample <- GeneratePoints(atp_serves[1:10,],arc3=10,plot=TRUE) 
 
ggplot() + geom_path(data = courtTrace, aes(x = x, y = y), color = 'black', size = 0.5) + 
  geom_segment(aes(x= 0, xend= 0, y= -6.5, yend= 6.5), size = 0.5, color = 'darkgrey', lineend = 'round') + 
  geom_line(aes(x=x,y=y, group=serveid), data=sample) 
 
ggplot() + geom_segment(aes(x=-11.89, xend=11.89, y=0, yend=0), size=0.5) +  
  geom_segment(aes(x=0,xend=0,y=0,yend=1.07), color='darkgrey') +  
  geom_line(aes(x=x, y=z, group=serveid, colour=time), data=sample) 
 
library(plotly) 
plot_ly(sample, x=x, y=y, z=z, group=serveid, type="scatter3d", mode="lines") %>% 
  add_trace(x=x, y=y, z=z, data=courtTrace, type="scatter3d", mode="lines") %>% 
  layout(scene=list(aspectmode="data")) 



#DEMO
### Court Trace ###
courtTrace <- data.frame(x = c(-11.89, -11.89, 0, 0, 0, 0, 0, 0, 11.89, 11.89, -11.89, -11.89, 11.89, 11.89, -11.89, -6.4, -6.4, 6.4, 6.4, 6.4, -6.4),
                         y = c(5.49, -5.49, -5.49, -6.5, -6.5, 6.5, 6.5, -5.49, -5.49, 5.49, 5.49, 4.115, 4.115, -4.115, -4.115, -4.115, 4.115, 4.115, -4.115, 0, 0),
                         z = c(0, 0, 0, 0, 1.09, 1.09, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))

#DEMO
sample <- GeneratePoints(atp_serves[1:10,], server, arc3=10,plot=TRUE)

ggplot() + geom_path(data = courtTrace, aes(x = x, y = y), color = 'black', size = 0.5) +
  geom_segment(aes(x= 0, xend= 0, y= -6.5, yend= 6.5), size = 0.5, color = 'darkgrey', lineend = 'round') +
  geom_line(aes(x=x,y=y, group=serveid), data=sample)

ggplot() + geom_segment(aes(x=-11.89, xend=11.89, y=0, yend=0), size=0.5) + 
  geom_segment(aes(x=0,xend=0,y=0,yend=1.07), color='darkgrey') + 
  geom_line(aes(x=x, y=z, group=serveid, colour=time), data=sample)



library(plotly)
plot_ly(sample, x=x, y=y, z=z, group=serveid, type="scatter3d", mode="lines") %>%
  add_trace(x=x, y=y, z=z, data=courtTrace, type="scatter3d", mode="lines") %>%
  layout(scene=list(aspectmode="data"))
library(dplyr)
library(purrr)

```

### What Variables should be used for Cluster Analysis? ###


```{r}


```

```{r, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE, fig.width=6, fig.height=4}
# Function for Hierarchical Clustering, with set Cluster Variables #
### The function will return a data frame with the serve ID and the cluster it falls into so it can be easily merged back to the data set


library(dplyr)
cluster_func<- function(data, nclust=8){
  mydata<- data %>% select_("serveid", "server", "center.x", "center.y", "speed", "start.x", "start.y", "start.z", "net.clearance", "height_off_bounce", "angle.change", "side", "direction.change") %>% mutate_each(funs(scale), 3:11)
  mydata$side<- factor(mydata$side)
  
  dist<-dist(mydata[,3:11])
  clusters<- hclust(dist)
  clustercut<- cutree(clusters,nclust)
  clustercut <- as.numeric(as.character(factor(clustercut)))
  out<- cbind(mydata$serveid,clustercut)
  colnames(out)<- c("serveid", paste0("cluster",nclust))
  return(as.data.frame(out))
}


## Function for Dendogram and Cluster Stats ##
cluster_analyse<- function(data, pcaN=c(4,6)){
  mydata<- data %>% select_("serveid", "server", "center.x", "center.y", "speed", "start.x", "start.y", "start.z", "net.clearance", "height_off_bounce", "angle.change", "side", "direction.change") %>% mutate_each(funs(scale), 3:11)
  mydata$side<- factor(mydata$side)
  
  dist<-dist(mydata[,3:11])
  clusters<- hclust(dist)
  plot(clusters, hang= -1)
  
  library(fpc)
  cl.stats <- NULL
  
  for (k in 2:15) {
    cl <- cutree(clusters, k)
    x <- cluster.stats(dist, cl)
    cl.stats <- rbind(cl.stats, c(k, x$within.cluster.ss, x$wb.ratio, x$ch,
                                      x$pearsongamma, x$dunn, x$dunn2))
  }
  colnames(cl.stats) <- c("cl","within.cluster.ss","wb.ratio", "ch", "pearsongamma", "dunn", "dunn2")
  cl.stats <- data.frame(cl.stats)
  cl.stats.m <- gather(cl.stats, key=stat, value=value, -cl)
  print(qplot(cl, value, data=cl.stats.m) + geom_line() + xlab("# clusters") + ylab("") +
  facet_wrap(~stat, ncol=3, scales = "free_y") + theme_bw())
  
  #PCA
  mydata_pca_val<- prcomp(mydata[,3:11], retx=T)
  mydata_pca<- NULL
  for (i in pcaN){
    new_pca_clust<- as.data.frame(cbind((mydata_pca_val$x), clust = cutree(clusters,i), nclust= i))
    mydata_pca<- rbind(mydata_pca, new_pca_clust)
  }
  ggplot(mydata_pca, aes(x=PC1, y=PC2, color=factor(clust))) + geom_point() + facet_wrap(~nclust)
}

```


### Cluster Analysis on first serves ###
```{r}
# Excluding Faults #
dat1<- atp_serves %>% filter(serve.x=="1") %>% filter(serve_classification < 3)
cluster_analyse(dat1, pcaN = c(4,6,8,10,12,14))
clust1.12<- cluster_func(dat1, 12)
atp_serves1.12<- merge(atp_serves,clust1.12, by="serveid")
```

## What is causing the outliers? Are these serves? What do they look like?

```{r}

#What is causing the outliers? ##
library(GGally)
clust1.4<- cluster_func(dat1, 4)
atp_serves1.4<- merge(atp_serves,clust1.4, by="serveid")
## Variable Split
ggscatmat(atp_serves1.4, columns = c("center.x", "center.y", "speed", "start.y", "height_off_bounce", "angle.change"), color= "cluster4")
## What do these odd clusters look like
ggplot(atp_serves1.4, aes(x=center.x, y=center.y, color=cluster4)) + geom_point() 

# Filtering for cluster 3
atp_serves1.4.3<- filter(atp_serves1.4, cluster4 == "3")
cluster1.4.3<- GeneratePoints(atp_serves1.4.3, server, arc3=10,plot=TRUE)

library(plotly)
plot_ly(cluster1.4.3, x=x, y=y, z=z, group=serveid, type="scatter3d", mode="lines") %>%
  add_trace(x=x, y=y, z=z, data=courtTrace, type="scatter3d", mode="lines") %>%
  layout(scene=list(aspectmode="data"))

```

Noticable outliers in the PCA were analysed deeper to find outliers effecting the clustering. When plotted in 3D, it is evident that the 2nd cluster in this 4 cluster solution are not serves, hence why speed and bounce was seen seperating the serves in the scatmat output. The 3rd cluster was notably a different type of serve that the spin variable, angle direction, and height of the bounce was having a large effect. When observing these serves they are seem to be kick serves that are to the Ad court side. Interestingly, these are all viable serves and the majority coming from Dominic Thiem (7 out of 17). These will be compared to second serves to see if this type of serve is more likely to fit into a second type of serve.  




## Profiling a 'n' Cluster Solution ### 

Variables for profiling
  - Average Speed
  - Variance of Speed
  - Spin
  - Average Arc
  - Side
  - Location (Wide, Body, T)
  - Servers 
  
  
```{r}
# Cluster Profiling - 12 Cluster Solution #
clust1.12<- cluster_func(dat1, 12)
clust1.12$cluster12 <- as.numeric(as.character(clust1.12$cluster12))
atp_serves1.12<- merge(atp_serves,clust1.12, by="serveid")

# Players in Each # 
table(atp_serves1.12$server, atp_serves1.12$cluster12)

# Cluster 8 ~ The Isner & Raonic Power Serve #
fast<- filter(atp_serves1.12, cluster12=="8")
fast<- GeneratePoints(fast, server, serve_classification, speed, arc3=10,plot=TRUE)

plot_ly(fast, x=x, y=y, z=z, group=serveid, type="scatter3d", mode="lines") %>%
  add_trace(x=x, y=y, z=z, data=courtTrace, type="scatter3d", mode="lines") %>%
  layout(scene=list(aspectmode="data"))

ggplot() + geom_path(data = courtTrace, aes(x = x, y = y), color= 'Black', size = 0.5) +
  geom_segment(aes(x= 0, xend= 0, y= -6.5, yend= 6.5), size = 0.5, color = 'darkgrey', lineend = 'round') +
  geom_line(aes(x=x,y=y, group=serveid, color=speed), data=fast) 

# Cluster 8 doesnt look visually different until you compare speeds

qplot(ordered(cluster12), speed, data=atp_serves1.12, geom="boxplot", color= ordered(cluster12))
table(atp_serves1.12$cluster12, atp_serves1.12$serve_classification)
qplot(factor(cluster12), height_off_bounce, data=atp_serves1.12, geom="boxplot", color= factor(cluster12))

# Height of the cluster 8 power serve starting position is a clear factor for the talls who can deliver these serves
qplot(factor(cluster12), start.z, data=atp_serves1.12, geom="boxplot", color= factor(cluster12))

# Ball locations
ggplot(atp_serves1.12, aes(x=center.x, y=center.y, color= factor(cluster12))) + geom_point() + facet_grid(~side)

```


## Profile Plots Combined ##
```{r, echo= FALSE}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

```

```{r}
# Summary Plots

library(ggplot2)
sumplots1<- qplot(factor(cluster12), speed, data=atp_serves1.12, geom="boxplot", color= factor(cluster12))
sumplots2<- qplot(factor(cluster12), height_off_bounce, data=atp_serves1.12, geom="boxplot", color= factor(cluster12))
sumplots3<- ggplot(atp_serves1.12, aes(x=center.x, y=center.y, color= factor(cluster12))) + geom_point() + facet_grid(~side)
sumplots4<- ggplot(atp_serves1.12, aes(x=start.y, y=start.z, color= factor(cluster12))) + geom_point()
sumplots5<- ggplot(atp_serves1.12, aes(x=center.y, y=height_off_bounce, color=factor(cluster12))) + geom_point()
sumplots6<- ggplot(atp_serves1.12, aes(x=center.y, y=net.clearance, color=factor(cluster12))) + geom_point()

multiplot(sumplots1, sumplots2, sumplots3, sumplots4, sumplots5, sumplots6, cols=2)
```



## Second Serve ##

```{r}
# Filter for second serve
dat2<- atp_serves %>% filter(serve.x=="2") %>% filter(serve_classification < 3)
# Analyse clusters
cluster_analyse(dat2, pcaN = c(3,4,6,7))
serve_clust2.7<- cluster_func(dat, 7)
# Merge with atp
atp_serves2.7<- merge(atp_serves,serve_clust2.7, by="serveid")
table(atp_serves2.7$cluster7)

library(GGally)
ggscatmat(atp_serves2.7, columns = c("center.x", "center.y", "speed", "start.y", "height_off_bounce", "angle.change"), color= "cluster7")


```

Observing the dendogram and the cluster stats plot, 4 or 7 clusters look to be present in the data. 

```{r}
# Cluster 1 and 7 could be an odd cluster
# What does cluster 1 look like?
atp_serves2.7.1<- filter(atp_serves2.7, cluster7 == "1")
cluster2.7.1<- GeneratePoints(atp_serves2.7.1, server, arc3=10,plot=TRUE)
table(atp_serves2.7.1$server)

require(plotly)
plot_ly(cluster2.7.1, x=x, y=y, z=z, group=serveid, type="scatter3d", mode="lines") %>%
  add_trace(x=x, y=y, z=z, data=courtTrace, type="scatter3d", mode="lines") %>%
  layout(scene=list(aspectmode="data"))

# What does cluster 2 look like?
atp_serves2.7.7<- filter(atp_serves2.7, cluster7 == "7")
cluster2.7.7<- GeneratePoints(atp_serves2.7.7, server, serve_classification, arc3=10,plot=TRUE)

require(plotly)
plot_ly(cluster2.7.7, x=x, y=y, z=z, group=serveid, type="scatter3d", mode="lines") %>%
  add_trace(x=x, y=y, z=z, data=courtTrace, type="scatter3d", mode="lines") %>%
  layout(scene=list(aspectmode="data"))

ggplot() + geom_path(data = courtTrace, aes(x = x, y = y), color= 'Black', size = 0.5) +
  geom_segment(aes(x= 0, xend= 0, y= -6.5, yend= 6.5), size = 0.5, color = 'darkgrey', lineend = 'round') +
  geom_line(aes(x=x,y=y, group=serveid, color=factor(serve_classification)), data=cluster2.7.7) 

# 2nd serve aces nothing to unusual, accept interesting to see three aces in this cluster. When compared to the other second serves this is quite interesting. The ggscatmat above shows the higher speed of these serves. 
table(atp_serves2.7$cluster7, atp_serves2.7$serve_classification)
```

```{r}

sumplots1.2<- qplot(factor(cluster7), speed, data=atp_serves2.7, geom="boxplot", color= factor(cluster7))
sumplots2.2<- qplot(factor(cluster7), height_off_bounce, data=atp_serves2.7, geom="boxplot", color= factor(cluster7))
sumplots3.2<- ggplot(atp_serves2.7, aes(x=center.x, y=center.y, color= factor(cluster7))) + geom_point() + facet_grid(~side)
sumplots4.2<- ggplot(atp_serves2.7, aes(x=start.y, y=start.z, color= factor(cluster7))) + geom_point()
sumplots5.2<- ggplot(atp_serves2.7, aes(x=center.y, y=height_off_bounce, color=factor(cluster7))) + geom_point()
sumplots6.2<- ggplot(atp_serves2.7, aes(x=center.y, y=net.clearance, color=factor(cluster7))) + geom_point()
 

multiplot(sumplots1.2, sumplots2.2, sumplots3.2, sumplots4.2, sumplots5.2, sumplots6.2, cols=2)

# Spin of second serves #
ggplot(atp_serves2.7, aes(x=speed, y=angle.change, color=factor(cluster7))) + geom_point()
ggplot(atp_serves2.7, aes(x=height_off_bounce, y=angle.change, color=factor(cluster7))) + geom_point()

```
 



```{r}

```





```{r}
## Comparing Raonic and Federer's serves ##
library(MASS)
select <- dplyr::select
Rao<- filter(atp_serves12, server=="RAONIC")
Fed<- filter(atp_serves12, server=="FEDERER")
Rao.clust<- table(Rao$clust12)
Rao.clust.relfreq <- Rao.clust / nrow(Rao)
barplot(Rao.clust.relfreq, ylim=c(0,0.4), main= "Raonic Frequency 1st Serves by cluster", col = "Light Blue")
Fed.clust<- table(Fed$clust12)
Fed.clust.relfreq <- Fed.clust / nrow(Fed) 
barplot(Fed.clust.relfreq, ylim=c(0,.6), main = " Federer Frequency 1st Serves by cluster", col = "Light Green")

```



```{r}
### Feature Generation Testing ###
# Height of bounce # 
summary(atp_serves$height_off_bounce)
hist(atp_serves$height_off_bounce)

# Direction Change #
# Plotting the the serves with changes in direction #
summary(atp_serves$direction.change)
summary(atp_serves$angle.change)
hist(atp_serves$angle.change)
changedirection<- atp_serves %>% filter(direction.change == "TRUE")
table(changedirection$serve_classification, factor(changedirection$server))


sample <- GeneratePoints(changedirection, server, side, server.x, serve_classification, arc3=10,plot=TRUE)

ggplot() + geom_path(data = courtTrace, aes(x = x, y = y), color = 'black', size = 0.5) +
  geom_segment(aes(x= 0, xend= 0, y= -6.5, yend= 6.5), size = 0.5, color = 'darkgrey', lineend = 'round') + geom_line(aes(x=x,y=y, group=serveid, color=server), data=sample)

ggplot() + geom_segment(aes(x=-11.89, xend=11.89, y=0, yend=0), size=0.5) + 
  geom_segment(aes(x=0,xend=0,y=0,yend=1.07), color='darkgrey') + 
  geom_line(aes(x=x, y=z, group=serveid, colour=time), data=sample)
## 3D Plot ##
library(plotly)
plot_ly(sample, x=x, y=y, z=z, group=serveid, type="scatter3d", mode="lines") %>%
  add_trace(x=x, y=y, z=z, data=courtTrace, type="scatter3d", mode="lines") %>%
  layout(scene=list(aspectmode="data"))

## Exploring serves with little or positive direction change ##
posdirectionchange<- atp_serves %>% filter(direction.change == "FALSE")
dcrandsample<- posdirectionchange[sample(nrow(posdirectionchange), 60), ]

sample2 <- GeneratePoints(dcrandsample, server, side, server.x, angle.change, serve_classification, arc3=10,plot=TRUE)

ggplot() + geom_path(data = courtTrace, aes(x = x, y = y), color = 'black', size = 0.5) +
  geom_segment(aes(x= 0, xend= 0, y= -6.5, yend= 6.5), size = 0.5, color = 'darkgrey', lineend = 'round') +
  geom_line(aes(x=x,y=y, group=serveid, color=angle.change), data=sample2)

library(plotly)
plot_ly(sample2, x=x, y=y, z=z, group=serveid, type="scatter3d", mode="lines") %>% add_trace(x=x, y=y, z=z, data=courtTrace, type="scatter3d", mode="lines") %>% layout(scene=list(aspectmode="data"))


```

*Outlier's Identification*
Unusual serve starting positions
Possible Flip
  Serve ID: 2_04_01_1_172646.trj, 2_04_01_1_172646.trj, 

```{r}
# Thomas Berdych Serve with a starting point 40cm+ higher than the second highest contact point.
tail(sort(atp_serves$start.z),5)
summary(atp_serves$start.x)
## Wrong starting Position ##
outlier<- GeneratePoints(atp_serves[c(481,1033),], server, side, server.x, angle.change, serve_classification, arc3=10,plot=TRUE)
plot_ly(outlier, x=x, y=y, z=z, group=serveid, type="scatter3d", mode="lines") %>% add_trace(x=x, y=y, z=z, data=courtTrace, type="scatter3d", mode="lines") %>% layout(scene=list(aspectmode="data"))

## NOT SERVES ###
head(sort(atp_serves$start.z))

## Identified through cluster analysis above (see cluster 2 in a four cluster solution), not actually serves, thought it was speed upon inspections speed is a consequence as they are ground shots after seeing starting height.

# Faults #


```
** Theory on different types of serves **
http://www.tennis4you.com/lesson-lounge/tennis4you/serve/4-different-serves.htm
https://www.coachup.com/resources/tennis/tennis-different-serving-methods?utm_campaign=client_organic&utm_content=Tennis_resource_DIFFERENT_TYPES_OF_TENNIS_SERVES&utm_medium=Active&utm_source=PR

4 types of Serves
Flat Serve
Slice Serve
Top-Spin Serve
Kick Serve


```



