---
title: "Playing Style"
author: "Braden Churcher"
date: "27 June 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load("~/Documents/UNI/Winter Research Project/Data/atp_serves.RData")

library(ggplot2)
library(dplyr)

library(purrr)


load("atp_serves.RData")


g <- function(coef, t) coef %*% t^(0:3)
g(unlist(atp_serves[1,c("raw.x0.1","raw.x1.1","raw.x2.1","raw.x3.1")]),atp_serves$start.1[1])

fedvmurray <- atp_serves[atp_serves$matchid=="AO2016F_Djokovic_Murray",]

table(atp_serves$server)
table(atp_serves$receiver)

```

## R Markdown
*Brain storm of serve styles*
List of Serve Characteristics for feature selection:
1st vs 2nd Serve
Points in game? if reciever is on 40 or server is on 40. Does this differ? are there risk takers?
Winning serves
Receivers Hand (R/L)
Riskiness and 
Types and key points:
Kick Serve:
  - Height of bounce (Z at )
      - Where was the Ball received?
  - Height over the net
  - Location ball is struck (is this different? To get top spin is it in a different spot)
  - Second serve characteristics
  - Speed
  - Spin? 
    - squared coefficients c2 --generate points function
    - How do we calculate this?
  - Location in Service Box

Down the tee (Flat vs Spin/Kick)
  - Ball location in service box, depth looks important
  - Speed
  - Location ball is hit. For right handers in the ad court this could be back past the centre line on the      deuce side
Outwide (Flat vs Spin/Kick)

Variables needed for Clustering:
Player - Server
X, Z, Y From start to finish.
Coefficients 
Speed
Receiver
Volicity 
Serve
Serve classification


** Random Plots ***  
## Roger Federer Receiving ##
```{r}

fedr <- atp_serves %>% filter(receiver == "FEDERER")
ggplot(fedr, aes(x=center.x, y=center.y, color=factor(serve_classification))) + geom_point() + facet_wrap(~side)

```
## Serve location by player ##
```{r}
ggplot(atp_serves_sub, aes(x=start.y, y=start.z, color=factor(server))) + geom_point()
```

## Most Aces ##

```{r}
table(atp_serves$server, atp_serves$serve_classification== "0")
```
## Raonic Aces ##
``` {r}
Rao<- atp_serves %>% filter(server == "RAONIC")
ggplot(Rao, aes(x=start.y, y=start.z, color=factor(serve_classification == "0"))) + geom_point()
ggplot(Rao, aes(x=start.x, y=start.z, color=factor(serve_classification == "0"))) + geom_point() 
ggplot(Rao, aes(x=center.x, y=center.y, color=factor(serve_classification))) + geom_point() 
```

``` {r}
# Net Clearance Height by 1st and second serve #
#work out how to solve for x in arc one at x=0 for all serves#
ggplot(atp_serves, aes(x=center.x, y=net.clearance, color=factor(serve.x))) + geom_point() 
ggplot(atp_serves, aes(x=center.y, y=net.clearance, color=factor(serve.x))) + geom_point() 
```
Using the distance at which the ball bounced in the x direction, you can see the difference in serves. The first serve being clearly longer and slightly lower over the net. 
Using the width at which the ball bounced in the court you can see the how second serves are generally much safer. The height over the net though may not be exactly easy to determine if it is different, there does seem to be less serves at the lower end of the scale and more variation in the second serve. 
Interestingly you can see how the net slouches in the the middle of the court. 

```{r}
# Different points in games #
ggplot(atp_serves, aes(x=center.x, y=center.y, color=factor(serve.x))) + geom_point() + facet_wrap(~receiver_score)

```


```{r}
# Serve classification by first and second serve #
table(atp_serves$serve_classification, atp_serves$serve.x)
```

Note that there are 37 double faults in the data, 6 aces on the second serve and when filtering by first serve for clustering, 357 serves will be lost. 

```{r}
## Roger Federer Receiving ##
fedr <- atp_serves %>% filter(receiver == "FEDERER")
ggplot(fedr, aes(x=center.x, y=center.y, color=factor(serve_classification))) + geom_point() + facet_wrap(~side)

## Most Aces ##
table(atp_serves$server, atp_serves$serve_classification== "0")

## Raonic Aces ##
Rao<- atp_serves %>% filter(server == "RAONIC")
ggplot(Rao, aes(x=start.y, y=start.z, color=factor(serve_classification == "0"))) + geom_point()
```


You can also embed plots, for example:

```{r pressure, echo=FALSE}
### Mitchell code for Court Position of Ball ###
traj_coords <- function(start, dir, flip, c0, c1, c2, c3, tm=0) {
  tm <- tm + start
  if (dir != "z") {
    c0 <- flip * c0
    c1 <- flip * c1
    c2 <- flip * c2
    c3 <- flip * c3
  }
  pos <- c0 + c1*tm + c2*tm^2 + c3*tm^3
  return(pos)
}

GeneratePoints <- function(data, ..., arc1=10, arc3=0, plot=FALSE){
  require(tidyr)
  require(dplyr)
  require(purrr)
  
  #arc2*(start.3 - (start.1+duration.arc1))
  data <- data %>% mutate(duration = arc1*duration.arc1 + arc3*duration.arc3, flip=sign(raw.x0.1))
  
  dots <- c("serveid", sapply(substitute(list(...))[-1], deparse))
  extravars <- match(dots[dots%in%colnames(data)], colnames(data))
  
  out <- data.frame()
  
  if(arc1!=0){
    arc1p <- data %>% select(extravars, start.1, duration.arc1, flip, raw.x0.1:raw.z3.1) %>%
      gather(coef, value, -(1:length(extravars)), -start.1, -duration.arc1, -flip) %>%
      separate(coef, c("junk1", "coef", "junk2"), sep="\\.") %>%
      select(-junk1, -junk2) %>%
      mutate(dir=substr(coef, 1, 1), coef=substr(coef, 2, 2)) %>%
      spread(coef, value) %>%
      rename("start" = `start.1`, "duration"=`duration.arc1`, "c0"=`0`, "c1"=`1`, "c2"=`2`, "c3"=`3`)
      arc1points <- seq(0,1,length.out=arc1) %>%
        map_df(function(x) mutate(rowwise(arc1p), time=start+duration*x, pos=traj_coords(start, dir, flip, c0, c1, c2, c3, tm=duration*x)))
      out <- rbind(out, cbind(arc=1, arc1points))
  }
  
  if(arc3!=0){
    arc3p <- data %>% select(extravars, start.3, duration.arc3, flip, raw.x0.3:raw.z3.3) %>%
      gather(coef, value, -(1:length(extravars)), -start.3, -duration.arc3, -flip) %>%
      separate(coef, c("junk1", "coef", "junk2"), sep="\\.") %>%
      select(-junk1, -junk2) %>%
      mutate(dir=substr(coef, 1, 1), coef=substr(coef, 2, 2)) %>%
      spread(coef, value) %>%
      rename("start" = `start.3`, "duration"=`duration.arc3`, "c0"=`0`, "c1"=`1`, "c2"=`2`, "c3"=`3`)
    arc3points <- seq(0,1, length.out=arc3) %>%
      map_df(function(x) mutate(rowwise(arc3p), time=start+duration*x, pos=traj_coords(start, dir, flip, c0, c1, c2, c3, tm=duration*x)))
    out <- rbind(out, cbind(arc=3, arc3points))
  }
  
  if(plot){
    extravars <- match(dots[dots%in%colnames(out)], colnames(out))
    out <- out %>% select(arc, extravars, flip, dir, time, pos) %>%
      spread(dir, pos)
  }
  
  return(out)
}

#DEMO
### Court Trace ###
courtTrace <- data.frame(x = c(-11.89, -11.89, 0, 0, 0, 0, 0, 0, 11.89, 11.89, -11.89, -11.89, 11.89, 11.89, -11.89, -6.4, -6.4, 6.4, 6.4, 6.4, -6.4),
                         y = c(5.49, -5.49, -5.49, -6.5, -6.5, 6.5, 6.5, -5.49, -5.49, 5.49, 5.49, 4.115, 4.115, -4.115, -4.115, -4.115, 4.115, 4.115, -4.115, 0, 0),
                         z = c(0, 0, 0, 0, 1.09, 1.09, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))

#DEMO
sample <- GeneratePoints(atp_serves[1:10,], server, arc3=10,plot=TRUE)

ggplot() + geom_path(data = courtTrace, aes(x = x, y = y), color = 'black', size = 0.5) +
  geom_segment(aes(x= 0, xend= 0, y= -6.5, yend= 6.5), size = 0.5, color = 'darkgrey', lineend = 'round') +
  geom_line(aes(x=x,y=y, group=serveid), data=sample)

ggplot() + geom_segment(aes(x=-11.89, xend=11.89, y=0, yend=0), size=0.5) + 
  geom_segment(aes(x=0,xend=0,y=0,yend=1.07), color='darkgrey') + 
  geom_line(aes(x=x, y=z, group=serveid, colour=time), data=sample)



library(plotly)
plot_ly(sample, x=x, y=y, z=z, group=serveid, type="scatter3d", mode="lines") %>%
  add_trace(x=x, y=y, z=z, data=courtTrace, type="scatter3d", mode="lines") %>%
  layout(scene=list(aspectmode="data"))
library(dplyr)
library(purrr)

```

### What Variables should be used for Cluster Analysis? ###

``` {r}
## Practice Cluster ##
# Cluster Second Coefficients #
test<- cbind(cluster=kmeans(atp_serves[,c("raw.x2.1", "raw.y2.1")], centers=2)$cluster, atp_serves)
ggplot(test, aes(x=raw.x2.1, y=raw.y2.1, colour=cluster)) + geom_point() + facet_wrap(~side)
ggplot(atp_serves, aes(x=x2.1, y=y2.1)) + geom_point() + facet_wrap(~side)
# note: Coefficents looked to be flipped. Does not seem to give anything usefullooking at these coefficients. 

```
# Creating Sample for Hierarchical Cluster Analysis #
First cluster analysis will just look at using a few different variables. Due to the differences in right handers serving to Ad and Deuce courts, the sample is split by the variable side to see what clusters are present on each side of the court (Ad=943, Deuce = 1057). These results will be compared later to see if the side of the court effects and gives rise to different clusters. 


```{r}
## First attempt ugly## 

## Cluster analysis on Ad Court side ##
adcourtclust<- atp_serves %>% filter(side=="Ad") %>% select(server, center.x, center.y, serve_classification, serve.x, start.x, start.y, start.z, net.clearance) 
adcourtclust$serve_classification<- factor(adcourtclust$serve_classification)
adcourtclust$serve.x<- factor(adcourtclust$serve.x)

clusters <- hclust(dist(adcourtclust[, 2:9]))
plot(clusters)
clustercut<- cutree(clusters,5)
adcourtclust$clust5<-clustercut
ggplot(adcourtclust, aes(x=center.x, y=center.y, color=factor(clust5))) + geom_point()

adcourtclustplot<- atp_serves %>% filter(side=="Ad") 
adcourtclustplot$clust5 <- clustercut
sampleclust5plot <- GeneratePoints(adcourtclustplot, clust5, server, arc3=10,plot=TRUE)
ggplot() + geom_path(data = courtTrace, aes(x = x, y = y), color= 'Black', size = 0.5) +
  geom_segment(aes(x= 0, xend= 0, y= -6.5, yend= 6.5), size = 0.5, color = 'darkgrey', lineend = 'round') +
  geom_line(aes(x=x,y=y, group=serveid, color=factor(clust5)), data=sampleclust5plot) 


## Deuce Court ##

deucecourtclust<- atp_serves %>% filter(side=="Deuce") %>% select(server, center.x, center.y, serve_classification, serve.x, start.x, start.y, start.z, net.clearance) 
deucecourtclust$serve_classification<- factor(deucecourtclust$serve_classification)
deucecourtclust$serve.x<- factor(deucecourtclust$serve.x)

clusters2 <- hclust(dist(deucecourtclust[, 2:9]))
plot(clusters2)
clustercut2<- cutree(clusters2,4)
deucecourtclust$clust4<-clustercut2
ggplot(deucecourtclust, aes(x=center.x, y=center.y, color=factor(clust4))) + geom_point()

deucecourtclustplot<- atp_serves %>% filter(side=="Deuce") 
deucecourtclustplot$clust4 <- clustercut2
deucesampleclust4plot <- GeneratePoints(deucecourtclustplot, clust4, server, arc3=10,plot=TRUE)
ggplot() + geom_path(data = courtTrace, aes(x = x, y = y), color= 'Black', size = 0.5) +
  geom_segment(aes(x= 0, xend= 0, y= -6.5, yend= 6.5), size = 0.5, color = 'darkgrey', lineend = 'round') +
  geom_line(aes(x=x,y=y, group=serveid, color=factor(clust4)), data=deucesampleclust4plot) 

```


```{r}

## Basic Hierarchical Cluster Using Given Variables ##
# Note using first serve only, filtering out faults #

# Sample for Clustering #
courtclust<- atp_serves %>% filter(serve.x=="1") %>% filter(serve_classification < 3) %>% select(serveid, server, center.x, center.y, speed, start.x, start.y, start.z, net.clearance, height_off_bounce, angle.change, side, direction.change) %>% mutate_each(funs(scale), 3:11)
courtclust$side<- factor(courtclust$side)

# Distance Calc #
dist<- dist(courtclust[,3:11])
clusters <- hclust(dist)
plot(clusters, hang= -1)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE, fig.width=6, fig.height=4}
# Cluster Stats #
library(fpc)
cl.stats <- NULL
for (k in 2:15) {
  cl <- cutree(clusters, k)
  x <- cluster.stats(dist, cl)
  cl.stats <- rbind(cl.stats, c(k, x$within.cluster.ss, x$wb.ratio, x$ch,
                                    x$pearsongamma, x$dunn, x$dunn2))
}
colnames(cl.stats) <- c("cl","within.cluster.ss","wb.ratio", "ch", "pearsongamma", "dunn", "dunn2")
cl.stats <- data.frame(cl.stats)
cl.stats.m <- gather(cl.stats, key=stat, value=value, -cl)
qplot(cl, value, data=cl.stats.m) + geom_line() + xlab("# clusters") + ylab("") +
  facet_wrap(~stat, ncol=3, scales = "free_y") + theme_bw()
```

```{r}
## Adding Different Cluster Solutions to Data for plotting ##
courtclustplot<- atp_serves %>% filter(serve.x=="1") %>% filter(serve_classification < 3) %>% select(serveid, server, center.x, center.y, speed, start.x, start.y, start.z, net.clearance, height_off_bounce, angle.change, side, direction.change) 
courtclustplot$side<- factor(courtclust$side)

# 3 Clusters
clustercut<- cutree(clusters,3)
courtclustplot$clust3<-clustercut
table(courtclustplot$clust3, courtclustplot$direction.change)

# 4 Clusters 
clustercut<- cutree(clusters,4)
courtclustplot$clust4<- clustercut


# 5 Clusters 
clustercut<- cutree(clusters,5)
courtclustplot$clust5<- clustercut
summary(factor(courtclustplot$clust5))

# 6 Clusters
clustercut<- cutree(clusters,6)
courtclustplot$clust6<- clustercut
summary(factor(courtclustplot$clust6))

# 7 Clusters
clustercut<- cutree(clusters,7)
courtclustplot$clust7<- clustercut
summary(factor(courtclustplot$clust7))

# 8 Clusters
clustercut<- cutree(clusters,8)
courtclustplot$clust8<- clustercut
summary(factor(courtclustplot$clust8))

# 9 Clusters
clustercut<- cutree(clusters,9)
courtclustplot$clust9<- clustercut
summary(factor(courtclustplot$clust9))

# 10 Clusters

clustercut<- cutree(clusters,10)
courtclustplot$clust10<- clustercut
summary(factor(courtclustplot$clust10))

# 11 Clusters

clustercut<- cutree(clusters,11)
courtclustplot$clust11<- clustercut
summary(factor(courtclustplot$clust11))

# 12 Clusters

clustercut<- cutree(clusters,12)
courtclustplot$clust12<- clustercut
summary(factor(courtclustplot$clust12))


ggplot(courtclustplot, aes(x=center.x, y=center.y, color=factor(clust11))) + geom_point() + facet_grid(~side)
```

```{r}
# Cluster Breakdown #

# PDA # 
courtclust_pca <- prcomp(courtclust[,3:11], retx=T)
plot(courtclust_pca, type= 'l')
courtclust_pca$rotation[,1:7]
courtclust_pca<- cbind((courtclust_pca$x),courtclustplot)
ggplot(courtclust_pca, aes(x=PC1, y=PC2, color=factor(clust12))) + geom_point()

#What is causing the outliers? ##
ggplot(courtclust_pca, aes(x=PC1, y=PC2, color=factor(clust4))) + geom_point()
library(GGally)
ggscatmat(courtclustplot, columns = c(3:5,7,10,11), color= "clust4")
courtclustplot$clust4<- factor(courtclustplot$clust4)

# Filtering for the 3rd cluster in a 4 cluster solution
cluster4.3<- filter(courtclustplot, clust4 == "3")
table(cluster4.3$server)
ggplot(cluster4.3, aes(x=center.x, y=center.y, color=factor(server))) + geom_point() 

cluster4.3plot<- select(cluster4.3, serveid)
cluster4.3plot<- merge(cluster4.3plot,atp_serves, by="serveid")
cluster4.3plot<- GeneratePoints(cluster4.3plot, server, center.x, center.y, arc3=10,plot=TRUE)

library(plotly)
plot_ly(cluster4.3plot, x=x, y=y, z=z, group=serveid, type="scatter3d", mode="lines") %>%
  add_trace(x=x, y=y, z=z, data=courtTrace, type="scatter3d", mode="lines") %>%
  layout(scene=list(aspectmode="data"))


# Filtering for the 2rd cluster in a 4 cluster solution #
cluster4.2<- filter(courtclustplot, clust4 == "2")
table(cluster4.2$server)
ggplot(cluster4.2, aes(x=center.x, y=center.y, color=factor(server))) + geom_point() 

cluster4.2plot<- select(cluster4.2, serveid)
cluster4.2plot<- merge(cluster4.2plot,atp_serves, by="serveid")
cluster4.2plot<- GeneratePoints(cluster4.2plot, server, center.x, center.y, arc3=10,plot=TRUE)

library(plotly)
plot_ly(cluster4.2plot, x=x, y=y, z=z, group=serveid, type="scatter3d", mode="lines") %>%
  add_trace(x=x, y=y, z=z, data=courtTrace, type="scatter3d", mode="lines") %>%
  layout(scene=list(aspectmode="data"))
```

For given serves working out comparable points is important. For simplicity on each polynomial we have decided to run with the variables given and add variables to account for spin and bounce height. 
Noticable outliers in the PCA were analysed deeper to find outliers effecting the clustering. To find out more about these serves, they were seperated and analysed. When plotted in 3D, it is evident that the 2nd cluster in this 4 cluster solution are not serves, hence why speed and bounce was seen seperating the serves in the scatmat output. The 3rd cluster was notably a different type of serve that the spin variable, angle direction, and height of the bounce was having a large effect. When observing these serves they are all kick serves that are to the Ad court side. Interestingly, these are all viable serves and the majority coming from Dominic Thiem (7 out of 17).



## Profiling a 'n' Cluster Solution ### 

Variables for profiling
  - Average Speed
  - Variance of Speed
  - Spin
  - Average Arc
  - Side
  - Location (Wide, Body, T)
  - Servers 
  
  
```{r}
# Cluster Profiling - 12 Cluster Solution #
atp_serves12<- merge(atp_serves,courtclustplot[,c("serveid","clust12")], by="serveid")

# Players in Each # 
table(atp_serves12$server, atp_serves12$clust12)

# Cluster 8 ~ The Isner & Raonic Power Serve #
fast<- filter(atp_serves12, clust12=="8")
fast<- GeneratePoints(Rao_Isn, server, serve_classification, speed, arc3=10,plot=TRUE)

plot_ly(fast, x=x, y=y, z=z, group=serveid, type="scatter3d", mode="lines") %>%
  add_trace(x=x, y=y, z=z, data=courtTrace, type="scatter3d", mode="lines") %>%
  layout(scene=list(aspectmode="data"))

ggplot() + geom_path(data = courtTrace, aes(x = x, y = y), color= 'Black', size = 0.5) +
  geom_segment(aes(x= 0, xend= 0, y= -6.5, yend= 6.5), size = 0.5, color = 'darkgrey', lineend = 'round') +
  geom_line(aes(x=x,y=y, group=serveid, color=speed), data=fast) 

# Cluster 8 doesnt look visually different until you compare speeds

qplot(factor(clust12), speed, data=atp_serves12, geom="boxplot", color= factor(clust12))
table(atp_serves12$clust12, atp_serves12$serve_classification)
qplot(factor(clust12), height_off_bounce, data=atp_serves12, geom="boxplot", color= factor(clust12))

# Height of the cluster 8 power serve starting position is a clear factor for the talls who can deliver these serves
qplot(factor(clust12), start.z, data=atp_serves12, geom="boxplot", color= factor(clust12))

# Ball locations
ggplot(atp_serves12, aes(x=center.x, y=center.y, color= factor(clust12))) + geom_point() + facet_grid(~side)

```


```{r}
## Comparing Raonic and Federer's serves ##
library(MASS)
Rao<- filter(atp_serves12, server=="RAONIC")
Fed<- filter(atp_serves12, server=="FEDERER")
Rao.clust<- table(Rao$clust12)
Rao.clust.relfreq <- Rao.clust / nrow(Rao)
barplot(Rao.clust.relfreq, ylim=c(0,0.4), main= "Raonic Frequency 1st Serves by cluster", col = "Light Blue")
Fed.clust<- table(Fed$clust12)
Fed.clust.relfreq <- Fed.clust / nrow(Fed) 
barplot(Fed.clust.relfreq, ylim=c(0,.6), main = " Federer Frequency 1st Serves by cluster", col = "Light Green")

```



```{r}
### Feature Generation Testing ###
# Height of bounce # 
summary(atp_serves$height_off_bounce)
hist(atp_serves$height_off_bounce)

# Direction Change #
# Plotting the the serves with changes in direction #
summary(atp_serves$direction.change)
summary(atp_serves$angle.change)
hist(atp_serves$angle.change)
changedirection<- atp_serves %>% filter(direction.change == "TRUE")
table(changedirection$serve_classification, factor(changedirection$server))


sample <- GeneratePoints(changedirection, server, side, server.x, serve_classification, arc3=10,plot=TRUE)

ggplot() + geom_path(data = courtTrace, aes(x = x, y = y), color = 'black', size = 0.5) +
  geom_segment(aes(x= 0, xend= 0, y= -6.5, yend= 6.5), size = 0.5, color = 'darkgrey', lineend = 'round') + geom_line(aes(x=x,y=y, group=serveid, color=server), data=sample)

ggplot() + geom_segment(aes(x=-11.89, xend=11.89, y=0, yend=0), size=0.5) + 
  geom_segment(aes(x=0,xend=0,y=0,yend=1.07), color='darkgrey') + 
  geom_line(aes(x=x, y=z, group=serveid, colour=time), data=sample)
## 3D Plot ##
library(plotly)
plot_ly(sample, x=x, y=y, z=z, group=serveid, type="scatter3d", mode="lines") %>%
  add_trace(x=x, y=y, z=z, data=courtTrace, type="scatter3d", mode="lines") %>%
  layout(scene=list(aspectmode="data"))

## Exploring serves with little or positive direction change ##
posdirectionchange<- atp_serves %>% filter(direction.change == "FALSE")
dcrandsample<- posdirectionchange[sample(nrow(posdirectionchange), 60), ]

sample2 <- GeneratePoints(dcrandsample, server, side, server.x, angle.change, serve_classification, arc3=10,plot=TRUE)

ggplot() + geom_path(data = courtTrace, aes(x = x, y = y), color = 'black', size = 0.5) +
  geom_segment(aes(x= 0, xend= 0, y= -6.5, yend= 6.5), size = 0.5, color = 'darkgrey', lineend = 'round') +
  geom_line(aes(x=x,y=y, group=serveid, color=angle.change), data=sample2)

library(plotly)
plot_ly(sample2, x=x, y=y, z=z, group=serveid, type="scatter3d", mode="lines") %>% add_trace(x=x, y=y, z=z, data=courtTrace, type="scatter3d", mode="lines") %>% layout(scene=list(aspectmode="data"))


```

*Outlier's Identification*
Unusual serve starting positions
Possible Flip
  Serve ID: 2_04_01_1_172646.trj, 2_04_01_1_172646.trj, 

```{r}
# Thomas Berdych Serve with a starting point 40cm+ higher than the second highest contact point.
tail(sort(atp_serves$start.z),5)
summary(atp_serves$start.x)
## Wrong starting Position ##
outlier<- GeneratePoints(atp_serves[c(481,1033),], server, side, server.x, angle.change, serve_classification, arc3=10,plot=TRUE)
plot_ly(outlier, x=x, y=y, z=z, group=serveid, type="scatter3d", mode="lines") %>% add_trace(x=x, y=y, z=z, data=courtTrace, type="scatter3d", mode="lines") %>% layout(scene=list(aspectmode="data"))

## NOT SERVES ###
head(sort(atp_serves$start.z))

## Identified through cluster analysis above (see cluster 2 in a four cluster solution), not actually serves, thought it was speed upon inspections speed is a consequence as they are ground shots after seeing starting height.

# Faults #
faults_serve


```
** Theory on different types of serves **
http://www.tennis4you.com/lesson-lounge/tennis4you/serve/4-different-serves.htm
https://www.coachup.com/resources/tennis/tennis-different-serving-methods?utm_campaign=client_organic&utm_content=Tennis_resource_DIFFERENT_TYPES_OF_TENNIS_SERVES&utm_medium=Active&utm_source=PR

4 types of Serves
Flat Serve
Slice Serve
Top-Spin Serve
Kick Serve


```



