---
title: "Classifying Playing Stlyes of Elite Tennis Players"
author: "Braden Churcher"
date: "10 July 2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

What is distinctive about the way Novak Djokovic or Roger Federer serves? What makes serves similar and what makes them different?

These questions refer to the 'playing style' of each player. We focus on the tennis serve due to it's importance in a tennis game as the first serve alone allows top male tennis players to win more than 80% of these points in a game.



(Johnson & McHugh, 2006)

## Overview of Hawkeye Data Set

The data set contains a sample of 2000 serves from the 2015 Australian Open Tournament.

- The serves are given by 2 Arc's, given by a function for the X,Y,Z position of the ball across time.
Arc 1: A Polynomial from the impact of the serve to when hits the ground. 
Arc 3: A Polynomial from where it hit the ground to where the ball either meets the opponents racket,           or end point if an ace. 

- Other data is also provided for each serve, ranging from server and receiver characterstics, speed, starting position, net clearance to the point characteristics or situation in the game when a serve is played. 


## What factors help distinguish Tennis Serves 

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

